The Hyperedge Event Model

### 2018-05-26 paper/ba-HEM/ba-sample.tex
new PPC plots

Concentrating on developing a statistical network model with better handling of hyperedges, we introduce the hyperedge event model (HEM) which jointly models the two components that govern time-stamped event formation: 1) the receiver selection processs that allows directed edges with one sender and one or more receivers or one receiver and one or more senders, and 2) the timing equation that models time increments (i.e., time to next event) with flexible distributional choices. In what follows, we introduce the HEM by describing how we assume the generatve process of a time-stamped event data (Section \ref{sec:generative process}), and deriving the conditional posteriors for Bayesian inference along with code review tests (Section \ref{sec:inference}). Then, we demonstrate the model's applicability by analyzing one of the county government email network---the Montgomery county email data---via model selection, posterior predictive checks, and exploratory analysis (Section \ref{subsec:Emails}). Finally, the paper finishes in Section 5 with conclusion and discussion.
\section{The Hyperedge Event Model}\label{sec:generative process}

Data generated under the model consists of $D$ unique edges. A single edge, indexed by $d \in [D]$, is represented by the three components: a sender $a_d \in [A]$, an indicator vector of receivers $\boldsymbol{r}_d = \{u_{dr} \}_{r=1}^{A}$, and a timestamp $t_d \in (0, \infty)$. For simplicity, we assume that edges are ordered by time such that $t_d \leq t_{d+1}$. While the model can be applied for two type of hyperedges---edges with (1) one sender and one or more receivers, and (2) one receiver and mone or more receivers---here we only present the generative process for those involving one sender and one or more receivers (i.e., multicast). For the latter case of hyperedges, we treat $a_d$ to be an indicator vector of senders $\boldsymbol{a}_d = \{u_{dr} \}_{a=1}^{A}$ and $r_d$ to be the single receiver. Detailed generative process for directed edges with one receiver and one or more senders are provided in Appendix A.

@@

@@

	Figure \ref{figure:PPCresults} illustrates the results of poterior predictive checks using log-normal distribution, which shows better performance in Section \ref{subsec:Experiment_email}. Upper two plots show node-specific posterior predictive degree distributions across $N=500$ synthetic data, where the left one for outdegree statistic and the right plot is for indegree statistic. For both plots, the x-axis represents the nodes ($A=1,\ldots,18$), and the y-axis represents the number of emails sent or received by the node. When compared with the observed outdegree and indegree statistics (red lines), our model recovers the overall distribution of sending and receiving activities across the nodes. For example, node 1 and 10 show significantly higher level of both sending and receiving activities relative to the rest, and the model-simulated data captures those big jumps, showing acceptable fit to the data. Outdegree distribution of some low-activity nodes are not precisely recovered, however, indegree distribution looks much better. Since we use more information in the receiver selection process (i.e., network effects) while we rely solely on minimum time increments when choosing the observed sender, these results are expected. Lower left plot is the distribution of receiver sizes, where the x-axis spans over the size of receivers 1 to 14 (which is the maximum size of observed receivers) and the y-axis denotes the number of emails with x-number of receivers. The result shows that our model is underestimating emails with one receiver while overestimating emails with two, three, and four receivers. One explanation behind what we observe is that the model is trying to recover so-called ``broadcast'' emails, which are the emails with $\geq 10$ number of receivers, so that the intercept estimate $b_1$ is slightly moved toward right. To improve the goodness of fit for receiver sizes, we can add more covariates (e.g., sender indicators) or assign strong prior structure on $b_1$. In the end, the plot on the lower right is the PP plot for time increments, which depicts the two cumulative distribution functions---one for simulated time increments and another for observed time increments---against each other in order to assess how closely two data sets agree. Here, the closeness to the diagonal line connecting $(0, 0)$ and $(1, 1)$ gives a measure of difference between the simulated and observed time increments, and our PP plot shows that we have great performance in reproducing the observed time distribution. All our findings from predictive experiments in Section \ref{subsec:Experiment_email} are further revealed in the PPC from exponential distribution, where the PPC plots comparing log-normal and exponential distributions are attached in Appendix C.
	\begin{figure}[!t]
		\centering
		\begin{tabular}[t]{cc}
		   	\begin{subfigure}[b]{0.495\textwidth}
		   		\caption{Outdegree distribution}
		\includegraphics[width=\textwidth]{img/outdegreenew-1.png}	
		   	\end{subfigure}
		   			   	\begin{subfigure}[b]{0.495\textwidth}
		   			   		\caption{Indegree distribution}
		   			   		\includegraphics[width=\textwidth]{img/indegreenew-1.png}	
		   			   	\end{subfigure}\\
			   	\begin{subfigure}[b]{0.495\textwidth}
			   		\caption{Receiver size distribution}
			   		\includegraphics[width=\textwidth]{img/recipientsizenew-1.png}	
			   	\end{subfigure}
			   	\begin{subfigure}[b]{0.495\textwidth}
			   				\centering
			   		\caption{PPplot for time increments}
			   		\includegraphics[width=0.6\textwidth]{img/timeppnew-1.png}	
			   	\end{subfigure}
		\end{tabular}
		\caption {PPC results from log-normal distribution. Blue lines denote the observed statistics in (a)--(c) and denotes the diagonal line in (d).}
		\label{figure:PPCresults}
	\end{figure}

@@

					\includegraphics[width=\textwidth]{img/outdegree_two-1.png}	

@@

					\includegraphics[width=\textwidth]{img/indegree_two-1.png}	

@@

					\includegraphics[width=\textwidth]{img/recipientsize_two-1.png}	

@@

					\includegraphics[width=0.625\textwidth]{img/timepp_two-1.png}

>>>>

Concentrating on developing a statistical network model with better handling of hyperedges, we introduce the hyperedge event model (HEM) which jointly models the two components that govern time-stamped event formation: 1) the receiver selection processs that allows directed edges with one sender and one or more receivers or one receiver and one or more senders, and 2) the timing equation that models time increments (i.e., time to next event) with flexible distributional choices. In what follows, we introduce the HEM by describing how we assume the generatve process of a time-stamped event data (Section \ref{sec:generative process}), and deriving the conditional posteriors for Bayesian inference along with code review tests (Section \ref{sec:inference}). Then, we demonstrate the model's applicability by analyzing one of the county government email network---the Montgomery county email data---via model selection, posterior predictive checks, and exploratory analysis (Section \ref{subsec:Emails}). Finally, the paper finishes in Section \ref{sec:conclusion} with conclusion and discussion.
\section{The Hyperedge Event Model}\label{sec:generative process}

Data generated under the model consists of $D$ unique edges. A single edge, indexed by $d \in [D]$, is represented by the three components: the sender $a_d \in [A]$, an indicator vector of recipients $\boldsymbol{r}_d = \{u_{dr} \}_{r=1}^{A}$, and the timestamp $t_d \in (0, \infty)$. For simplicity, we assume that edges are ordered by time such that $t_d \leq t_{d+1}$. While the model can be applied for two type of hyperedges---edges with (1) a single sender and multiple receivers, and (2) multiple senders and a single receiver---here we only present the generative process for those involving a single sender and multiple receivers (i.e., multicast). For the latter case of hyperedges, we treat $a_d$ to be an indicator vector of senders $\boldsymbol{a}_d = \{u_{dr} \}_{a=1}^{A}$ and $r_d$ to be the single recipient.

@@

										\includegraphics[width=0.4\textwidth]{img/modellabel.png}				

@@

	Figure \ref{figure:PPCresults} illustrates the results of poterior predictive checks using log-normal distribution, which shows better performance in Section \ref{subsec:Experiment_email}. Upper two plots show node-specific posterior predictive degree distributions across $N=500$ synthetic data, where the left one for outdegree statistic and the right plot is for indegree statistic. For both plots, the x-axis represents the nodes ($A=1,\ldots,18$), and the y-axis represents the number of emails sent or received by the node. When compared with the observed outdegree and indegree statistics (red lines), our model recovers the overall distribution of sending and receiving activities across the nodes. For example, node 1 and 10 show significantly higher level of both sending and receiving activities relative to the rest, and the model-simulated data captures those big jumps, showing acceptable fit to the data. Outdegree distribution of some low-activity nodes are not precisely recovered, however, indegree distribution looks much better. Since we use more information in the receiver selection process (i.e., network effects) while we rely solely on minimum time increments when choosing the observed sender, these results are expected. Lower left plot is the distribution of receiver sizes, where the x-axis spans over the size of receivers 1 to 14 (which is the maximum size of observed receivers) and the y-axis denotes the number of emails with x-number of receivers. The result shows that our model is underestimating emails with one receiver while overestimating emails with two, three, and four receivers. One explanation behind what we observe is that the model is trying to recover so-called ``broadcast'' emails, which are the emails with $\geq 10$ number of receivers, so that the intercept estimate $b_1$ is slightly moved toward right. To improve the goodness of fit for receiver sizes, we can add more covariates (e.g., sender indicators) or assign strong prior structure on $b_1$. In the end, the plot on the lower right is the PP plot for time increments, which depicts the two cumulative distribution functions---one for simulated time increments and another for observed time increments---against each other in order to assess how closely two data sets agree. Here, the closeness to the diagonal line connecting $(0, 0)$ and $(1, 1)$ gives a measure of difference between the simulated and observed time increments, and our PP plot shows that we have great performance in reproducing the observed time distribution. All our findings from predictive experiments in Section \ref{subsec:Experiment_email} are further revealed in the PPC from exponential distribution, where the PPC plots comparing log-normal and exponential distributions are attached in Appendix C.
	\begin{figure}[!t]
		\centering
		\begin{tabular}[t]{cc}
		   	\begin{subfigure}[b]{0.495\textwidth}
		   		\caption{Outdegree distribution}
		\includegraphics[width=\textwidth]{img/outdegree-1.png}	
		   	\end{subfigure}
		   			   	\begin{subfigure}[b]{0.495\textwidth}
		   			   		\caption{Indegree distribution}
		   			   		\includegraphics[width=\textwidth]{img/indegree-1.png}	
		   			   	\end{subfigure}\\
			   	\begin{subfigure}[b]{0.495\textwidth}
			   		\caption{Receiver size distribution}
			   		\includegraphics[width=\textwidth]{img/recipientsize-1.png}	
			   	\end{subfigure}
			   	\begin{subfigure}[b]{0.495\textwidth}
			   				\centering
			   		\caption{PPplot for time increments}
			   		\includegraphics[width=0.56\textwidth]{img/timePPplot-1.png}	
			   	\end{subfigure}
		\end{tabular}
		\caption {PPC results from log-normal distribution. Blue lines denote the observed statistics in (a)--(c) and denotes the diagonal line in (d).}
		\label{figure:PPCresults}
	\end{figure}

@@

					\includegraphics[width=\textwidth]{img/outdegree2-1.png}	

@@

					\includegraphics[width=\textwidth]{img/indegree2-1.png}	

@@

					\includegraphics[width=\textwidth]{img/recipientsize2-1.png}	

@@

					\includegraphics[width=0.56\textwidth]{img/timePPplot2-1.png}
